---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# scoccer

Easen the usage of scottish soccer data from data from http://www.football-data.co.uk

## Installation

You can install scoccer from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("steffenbank/scoccer")
```

## Basic usage

The most usefull functions within the packages are \code{sco_acquire} and \code{sco_relative_strength}. Description of fields are availble below @ the column notes. Currently are two leagues available:

1. Scottish Premiership ("sco_pl")
2. Scottish Champtionship ("sco_ch")

Years from 2001 till now are available - season inputs may be specificed as "yyyy", for instance 2018/2019 season "1819". 

Data from Scottish Premiership 2017/2018:
``` r
library(scoccer)
head(sco_acquire("1718", "sco_pl"))
```

Relative strength between Ayr and Partick Thistle the last five games. May be used for model purposes:
``` r
library(scoccer)
sco_relative_strength("1819","sco_pl","Ayr","Partick",5) 
```

## Modeling use

## Column notes
Key to results data:

Div = League Division \n
Date = Match Date (dd/mm/yy)
HomeTeam = Home Team
AwayTeam = Away Team
FTHG and HG = Full Time Home Team Goals
FTAG and AG = Full Time Away Team Goals
FTR and Res = Full Time Result (H=Home Win, D=Draw, A=Away Win)
HTHG = Half Time Home Team Goals
HTAG = Half Time Away Team Goals
HTR = Half Time Result (H=Home Win, D=Draw, A=Away Win)

Match Statistics (where available)
Attendance = Crowd Attendance
Referee = Match Referee
HS = Home Team Shots
AS = Away Team Shots
HST = Home Team Shots on Target
AST = Away Team Shots on Target
HHW = Home Team Hit Woodwork
AHW = Away Team Hit Woodwork
HC = Home Team Corners
AC = Away Team Corners
HF = Home Team Fouls Committed
AF = Away Team Fouls Committed
HFKC = Home Team Free Kicks Conceded
AFKC = Away Team Free Kicks Conceded
HO = Home Team Offsides
AO = Away Team Offsides
HY = Home Team Yellow Cards
AY = Away Team Yellow Cards
HR = Home Team Red Cards
AR = Away Team Red Cards
